Rising from the Ashes II: The Bar-driven Abundance Bimodality of the Milky Way

### 2024-10-11 paper/ms.tex
Charlie comments

\begin{abstract}
    The Milky Way hosts at least two modes in its present day distribution of Fe and $\alpha$-elements. The exact cause of this bimodality is disputed, but one class of explanations involves the merger between the Milky Way and a relatively massive satellite (Gaia-Sausage-Enceladus) at $z\sim2$. However, reproducing this bimodality in simulations is not straightforward, with conflicting results on the prevalance, morphology, and mechanism behind multimodality. We present a case study of a galaxy in the Illustris TNG50 simulation which undergoes a sequence of starburst, brief quiescence, and then rejuvenation.
    This sequence results in a pronounced abundance bimodality after a post-processing adjustment of the \alphaFe{} of old stars which is designed to mimic a higher star formation efficiency in dense gas.
    The high- and low-$\alpha$ sequences are neatly separated in time by the brief quiescent period. The quiescent period in this galaxy is not associated with a merger but by the formation of a bar followed by AGN activity. This galaxy indicates a novel scenario in which the $\alpha$-bimodality in the Milky Way is caused by AGN-induced quenching triggered by the formation of its bar. We predict that abundance bimodalities should be more common in barred as opposed to unbarred galaxies.
  \end{abstract}

@@

\section{Introduction}\label{sec:intro}
The stellar surface abundances of most elements retain the composition of their natal gas cloud. Therefore, the present-day distribution of their abundances encodes the chemical history of a galaxy's gas phase. Two classes of elements have received particular attention in the Milky Way due to their disparate formation channels: iron-peak (such as Fe) and $\alpha$-elements (produced through the $\alpha$-process, such as O and Mg). Iron-peak elements are produced in both Type~Ia and Type~II supernovae (SNe) whereas $\alpha$-elements are produced predominantly through Type~II SNe. In the Milky Way, there is a well-established bimodality with separate high- and low-$\alpha$ sequences shown in Figure~\ref{fig:fig1} \citep{1996ASPC...92..307G,1998A&A...338..161F,2004AN....325....3F,2006MNRAS.367.1329R,2011A&A...535L..11A,2012A&A...545A..32A,2014A&A...562A..71B,2014ApJ...796...38N,2020MNRAS.493.2952H}.

Broadly speaking, there are three approaches to explaining the bimodality. First, it is a result of internal secular processes that generate the bimodality through radial migration \citep{2009MNRAS.396..203S,2021MNRAS.507.5882S,2023MNRAS.523.3791C} or clump formation \citep{2019MNRAS.484.3476C,2020MNRAS.492.4716B,2021MNRAS.502..260B,2023ApJ...953..128G}. Second, the bimodality is generated through gas infall scenarios, either from specific gas accretion episodes from the intergalactic medium \citep{1997ApJ...477..765C,2009IAUS..254..191C,2017MNRAS.472.3637G,2019A&A...623A..60S}, or through a more self-consistent collapse sequence of the circumgalactic medium driven through feedback \citep{2021MNRAS.501.5176K}. Third and finally, the bimodality is generated through a merger process, either by enhancing the star formation rate (SFR) of the Galaxy \citep{2004ApJ...612..894B,2005ApJ...630..298B,2007ApJ...658...60B,2010MNRAS.402.1489R} or by supplying a relatively pristine gas supply that resets the metallicity of the Galaxy \citep{2020MNRAS.491.5435B,2024MNRAS.528L.122C}. Strong evidence that the Milky Way did undergo a merger with the Gaia-Sausage-Enceladus (GSE) satellite supports these scenarios \citep{2018MNRAS.478..611B,2018Natur.563...85H,2020ApJ...901...48N,2024ApJ...972..112C}.

@@

In Paper~I, we used idealized simulations of a galaxy merger that triggered the quiescent period. However, in that work we argued the merger was unnecessary. In this work, we study a subhalo, or galaxy, from the Illustris TNG50 simulation demonstrating this. This subhalo of interest (SoI) exhibits the sequence of events presented in Paper~I after a simple post-processing step which increases the \alphaFe{} of old star particles. This post-processing is motivated by recent work demonstrating that the star formation efficiency (SFE) of the dense gas present at high-$z$ is too low in the TNG model \citep{2024arXiv240909121H}. The SoI undergoes a brief quiescent period which neatly separates a high- and low-$\alpha$ sequence. However, instead of being preceded by a merger, the quenching is preceded by apparent bar-induced AGN activity. Therefore, this work serves as a verification that the scenario in Paper~I is possible in cosmological simulations and does not require a merger.

In Section~\ref{sec:methods}, we discuss our selection technique which led to discovering the SoI, the observations we use for comparison, as well as a simple one zone chemical evolution model we use to justify our post-processing step. In Section~\ref{sec:results}, we present the main results which we discuss and interpret in Section~\ref{sec:disc}. We conclude in Section~\ref{sec:conc}.

@@

We select subhalo 172175 (the ID at snapshot 40) for its resemblance to the Milky Way. We then studied the main descendant of this subhalo at $z=0$ (subhalo 392276 at snapshot 99). A summary of its key properties is given in Table~\ref{tab:summ}. We refer to this subhalo as our subhalo of interest (SoI).

@@

The SoI, on the other hand, does not show a clearly bimodal structure in the fiducial simulation (middle column). There is some structure in the $\FeH=-0.75$ bin. The right panel of Figure~\ref{fig:fig1} shows the same distribution as in the middle panel, but with a modification to increase \MgFe{} values of older stars formed before $z=1.5$ (see Sections~\ref{ssec:tng} and \ref{ssec:onezone}). A multimodal structure emerges with three clear modes at $\MgFe\sim0.8$, $0.5$, and $0.2\,\dex$. The 1D histograms show that the modes are well-separated, and that the troughs between the modes nearly vanish.

@@

\begin{figure*}
  \centering
  \includegraphics[width=\textwidth]{fig/392276_alpha.pdf}
  \caption{\textbf{Bimodality in the abundance plane is linked to distinct epochs in simulation.} The upper panels show \MgFe{} as a function of age for our subhalo in TNG. The colors indicate stellar populations at fixed values of \FeH{}, which are the same as in Figure~\ref{fig:fig1}. A gap in the relation occurs at an age of approximately $10.6\,\Gyr$, which we indicate with a vertical dashed line. The effect of the $\alpha$-enhancement is clear, as it separates the stars that form before and after this gap in ages (star particles which formed before $z=1.5$ are $\alpha$-enhanced, which occurs at an age of $\sim9.5\,\Gyr$). The lower panels show on the left the Milky Way and on the center and right the data from TNG but with $12.5\%$ age errors and $0.01\,\dex$ errors in \MgFe{}. When the simulations are given these errors, we see that the before and after star particles smear such that the two populations significantly overlap in ages. In the $\alpha$-enhanced SoI, two populations emerge in each bin which overlapped in the fiducial distribution. This feature more closely resembles the Milky Way, which displays such populations where the bimodality is strongest -- $\FeH=-0.5$ (blue) and $-0.25$ (orange).}
  \label{fig:alpha}
\end{figure*}

The abundance distributions shown in Figure~\ref{fig:fig1} can be better understood by examining the evolution of \MgFe{} with time of the individual stars/star particles. In the upper panels of Figure~\ref{fig:alpha} we show the true distribution of \MgFe{} as a function of time for the fiducial SoI in the middle and for the post-processed, $\alpha$-enhanced subhalo to the right. We use age instead of formation time in order to better facilitate comparisons to observations. These panels show 2D histograms, with a logarithmic colormap normalized to the maximum of the plot. To prevent overlap, the values of \MgFe{} are given offsets of $0$, $-0.5$, $-1$, and $-1.5$, in order of increasing \FeH{}.

In the SoI, there is an age gap at $\sim10.6\,\Gyr$, which we mark with a vertical dashed line in the upper row. Star particles older than this line have a much clearer gradient in time with \MgFe{} than stars that form after. In the $\FeH=-0.25$ bin, star particles which form directly after this line have a slightly reduced \MgFe{} than stars which form a short time later.

In the lower row, the left panel shows data from the Milky Way, while the middle and right panels correspond to the fiducial and $\alpha$-enhanced SoI, respectively. For the simulations, we introduce Gaussian errors of $12.5\%$ in age and $0.01,\dex$ in \MgFe{}. These error values align with the observational uncertainties from the APOKASC-3 and APOGEE datasets, which the Milky Way panel (lower left) is based on (see Appendix~\ref{app:obs_err} for detailed plots of observational errors). While the error in \MgFe{} is insignificant, the age error ($1.5\,\Gyr$ at $10\,\Gyr$) blurs the distribution, particularly around the dashed line that marks the transition between sequences. Despite this, the $\alpha$-enhanced SoI still shows two distinct populations, although their ages now overlap more significantly.

Interestingly, the Milky Way distribution (lower left panel) exhibits similarities to the $\alpha$-enhanced SoI. Specifically, in the $\FeH=-0.5$ and $-0.25$ bins (blue and orange, respectively), there appear to be two populations that overlap in age but remain distinct in \MgFe{}. These are the bins where the bimodality is most pronounced (as seen in the upper left panel of Figure~\ref{fig:fig1}).

@@

To understand the key events driving the behavior around the dashed line in Figure~\ref{fig:alpha}, we examine the evolution of several properties of the SoI in Figure~\ref{fig:history}. The vertical dashed line in each panel marks the transition at $t\sim3.2\,\Gyr$ from the high- to low-$\alpha$ sequences, as in Figure~\ref{fig:alpha}. The upper panel in the left column shows the the SFR (computed for all gas cells in the SoI). There are two peaks at $t\sim2.5\,\Gyr$ and $t\sim4.5\,\Gyr$ with maximum values of $50\,\Msun/\textrm{yr}$ and $30\,\Msun/\textrm{yr}$, respectively. Around the high- to low-$\alpha$ transition, there is a dip in the SFR, which drops by an order of magnitude to about $3\,\Msun/\textrm{yr}$. The right panel zooms in on the period between $t=2\,\Gyr$ and $5\,\Gyr$, where we observe a sharp recovery in the SFR following the quiescent phase. In the span of a single snapshot (roughly $150\,\Myr$), the SFR increases from about $3\,\Msun/\textrm{yr}$ to $10\,\Msun/\textrm{yr}$.

The middle panels track the accretion rate of the central BH as a fraction of the Eddington rate. Early in the SoI's history ($t<2\,\Gyr$), the BH experiences high accretion, which steadily declines until $t\sim5\,\Gyr$. Around $t\sim3.2\,\Gyr$, the BH accretion rate peaks again, reaching approximately $30\%$ of the Eddington limit, placing the BH in quasar mode and injecting significant thermal energy into the galaxy's center. The middle right panel shows the period between $t=2$ and $5\,\Gyr$.

The lower panel illustrates the growth of the SoI's virial mass ($M_{200}$). Early on ($t<4\,\Gyr$), $M_{200}$ increases roughly linearly, reaching about $2 \times 10^{12}\,\Msun$. After this, the mass remains relatively stable until jumps occur around $t\sim10$ and $\sim12\,\Gyr$, indicative of mergers. There are no mergers related to the earlier quiescent period around $t\sim3.2\,\Gyr$, as no major mass jumps are observed during this time. The lower right panel shows the period between $t=2$ and $5\,\Gyr$.

\subsection{Sequence of Events}\label{ssec:sequence_of_events}
\begin{figure*}
  \centering
  \includegraphics[width=\textwidth]{fig/alyssa.pdf}
  \caption{Surface density projections of gas (top row) and star particles (middle row) in our SoI across snapshots at different times during the high- to low-$\alpha$ transition. Below the projections is a plot showing the SFR, BH accretion rate (in units of $\dot{M}_{\textrm{BH}}/\dot{M}_{\textrm{edd}}$), and bar strength ($A_2/A_0$ for $R<2$ kpc). Time ranges from $\sim2.4\,\textrm{Gyr}$ to $\sim3.6\,\textrm{Gyr}$, corresponding to redshifts from $z\sim2.7$ to $z\sim1.8$. A sequence of events in which the bar strengthens, BH accretion increases, and SFR declines is seen, and is described more fully in the text.}
  \label{fig:seq}
\end{figure*}

@@

\section{Discussion}\label{sec:disc}
In Figure~\ref{fig:fig1}, we compared the abundance plane between the Milky Way and our SoI before and after our $\alpha$-enhancement post-processing procedure. It is clear that the TNG SoI is unimodal before the $\alpha$-enhancement and multi-modal afterwards. Here, we briefly discuss two main points: (1) assuming the $\alpha$-enhancement is justified, what leads to the bimodality in the SoI?, and (2) what justifies the $\alpha$-enhancement? We then extend our comparison to data and argue the quiescent period is driven by a bar-induced AGN episode.

@@

Once the $\alpha$-enhancement post-processing has been done, the SoI that we have studied in this work is consistent with this scenario. The vertical dashed line in Figures~\ref{fig:alpha} and \ref{fig:history} denotes the transition between the high- and low-$\alpha$ sequences. We place it at a lookback time of $10.6\,\Gyr$ or a cosmic time of $\sim3.2\,\Gyr$ or $z\sim2.05$.\footnote{The transition between the high- and low-$\alpha$ sequences occurs approximately $1\,\Gyr$ before $z=1.5$, where the $\alpha$-enhancement begins. This, combined with the fact that not all of the subhalos in our sample display bimodalities (Appendix~\ref{app:rand_fig1}), shows that the $\alpha$-enhancement is not the direct cause of the bimodality.} The upper right panel of Figure~\ref{fig:alpha} shows 2D histograms of \MgFe{} vs. star particle age in fixed bins of \FeH{} (each color is a different \FeH{} bin, and an offset is given to \MgFe{} to prevent overlap). Stars formed before the dashed line had a steep decline of \MgFe{} with time, while the relation in star particles formed after the dashed line is much more flat. At the $\alpha$-transition, we see a gap in the $\FeH=-0.5$ and $-0.25$ bins.

@@

It is also worth noting that in both the fiducial and $\alpha$-enhanced SoI show a slight rebound effect in \MgFe{}. The star particles which form directly after the dashed line have a slightly lower \MgFe{} than stars which form after. This was also predicted in Figure~9 of Paper~I, where it was argued that during the period of suppressed star formation, the \alphaFe{} ratio of the star-forming gas drops sharply due to the reduced contribution of Type~II relative to Type~Ia SNe. Later, the \alphaFe{} of the gas will recover when the SFR also recovers, but there is a brief window when old, low-$\alpha$ stars can form. A similar behavior was seen in the one-zone models with bursty SFHs in \citet{2020MNRAS.498.1364J}.

@@

\subsection{Direct Comparison to Observations}\label{ssec:compare_obs}
The lower left panel of Figure~\ref{fig:alpha} shows the \MgFe{} vs. age of stars in bins of \FeH{} which pass our solar neighborhood selection and are present in the APOKASC-3 catalog. In the bins where the bimodality is strongest (blue and orange, $\FeH=-0.5$ and $-0.25$, respectively), we see that there is a sort of two tiered distribution with significant overlap in age. In the SoI with age and abundance errors shown in the lower right panel, a similar distribution can be seen in all \FeH{} bins. An examination of the true distribution in the SoI (upper right panel), we see that the two distributions are in fact very cleanly separated in age.

@@

\subsection{Cause of Quiescence}\label{ssec:cause_qui}
In Paper~I, AGN activity from a merger was the suspected cause for the quiescent period. In our SoI, there is indeed a brief burst in AGN accretion at the time of the merger (middle panel of Figure~\ref{fig:history}). Based on this burst, it is also reasonable to suspect that AGN activity is also responsible for the quiescent period in our SoI. However, we argue that the cause of the localized spike in the BH accretion rate is not due to a merger but instead due to the formation/strengthening of a bar.

@@

A complicating factor for this picture comes from the high SFR associated with the galaxy. The depletion time ($M_{\textrm{gas}}/\textrm{SFR}$) at the $t=2.84\,\Gyr$ snapshot is $204\,\Myr$, shorter than the time it takes for the SFR to reach its minimum. This implies the possibility of starvation as a quenching mechanism. However, the average depletion time in the preceding 5 snapshots (which are each $\sim150\,\Myr$ apart) is $220\,\Myr$, so clearly the SoI is accreting high amounts of gas from its environment. A definitive account, difficult with the current simulation outputs because of its sparse snapshot spacing, requires further work.

\section{Conclusions}\label{sec:conc}
In this work, we examined a subhalo of interest in Illustris TNG50. This SoI is at a Milky Way-progenitor mass at $z=1.5$. After applying a post-processing step that increased the \MgFe{} of star particles formed before $z=1.5$, this subhalo hosts a strong bimodality in the plane of \MgFe{} and \FeH{}, shown in Figure~\ref{fig:fig1}. This post-processing is justified by arguing that the SFE of dense gas is too low in TNG \citep[][see discussion in our Section~\ref{ssec:sfe}]{2024arXiv240909121H}.

@@

\section{Random Selection of Subhalos}\label{app:rand_fig1}
In Figures~\ref{fig:app0} to \ref{fig:app16}, we show a version of Figure~\ref{fig:fig1} but for a random selection of subhalos from our broader sample of 168 Milky Way-progenitors at $z=1.5$ in TNG50. It is from this sample that we selected our SoI. We first show our SoI at $z=1.5$ in Figure~\ref{fig:app0}, and then show our random sample of 16.

\begin{figure*}
  \centering
  \includegraphics[width=\textwidth]{fig/app_172175.pdf}
  \caption{The same as Figure~\ref{fig:fig1}, but for our SoI at $z=1.5$.}
  \label{fig:app0}
\end{figure*}

>>>>

\begin{abstract}
    The Milky Way hosts at least two modes in its present day distribution of Fe and $\alpha$-elements. The exact cause of this bimodality is disputed, but one class of explanations involves the merger between the Milky Way and a relatively massive satellite (Gaia-Sausage-Enceladus) at $z\sim2$. However, reproducing this bimodality in simulations is not straightforward, with conflicting results on the prevalance, morphology, and mechanism behind multimodality. We present a case study of a galaxy in the Illustris TNG50 simulation which undergoes a sequence of starburst, brief quiescence, and then rejuvenation. This sequence results in a pronounced abundance bimodality after a post-processing adjustment of the \alphaFe{} of old stars which is designed to mimic a higher star formation efficiency in dense gas. The high- and low-$\alpha$ sequences are neatly separated in time by the brief quiescent period. The quiescent period in this galaxy is not associated with a merger but by the formation of a bar followed by AGN activity. This galaxy indicates a novel scenario in which the $\alpha$-bimodality in the Milky Way is caused by the formation of the bar via AGN-induced quenching. We predict that abundance bimodalities should be more common in barred as opposed to unbarred galaxies.
  \end{abstract}

@@

\section{Introduction}\label{sec:intro}
The stellar surface abundances of most elements retain the composition of their natal gas cloud. Therefore, the present-day distribution of their abundances encodes the chemical history of a galaxy's gas phase. Two classes of elements have received particular attention in the Milky Way due to their disparate formation channels: iron-peak (such as Fe) and $\alpha$-elements (produced through the $\alpha$-process, such as O and Mg). Enrichment of iron-peak elements is primarily through Type~Ia and Type~II supernovae (SNe), whereas enrichment of $\alpha$-elements is primarily only through Type~II SNe. In the Milky Way, there is a well-established bimodality with separate high- and low-$\alpha$ sequences shown in Figure~\ref{fig:fig1} \citep{1996ASPC...92..307G,1998A&A...338..161F,2004AN....325....3F,2006MNRAS.367.1329R,2011A&A...535L..11A,2012A&A...545A..32A,2014A&A...562A..71B,2014ApJ...796...38N,2020MNRAS.493.2952H}.

Broadly speaking, there are three approaches to explaining the bimodality. First, it is a result of internal secular processes that generate the bimodality through radial migration \citep{2009MNRAS.396..203S,2021MNRAS.507.5882S,2023MNRAS.523.3791C} or clump formation \citep{2019MNRAS.484.3476C,2020MNRAS.492.4716B,2021MNRAS.502..260B,2023ApJ...953..128G}. Second, the bimodality is generated through gas infall scenarios, either from specific gas accretion episodes from the intergalactic medium \citep{1997ApJ...477..765C,2009IAUS..254..191C,2017MNRAS.472.3637G,2019A&A...623A..60S}, or through a more self-consistent collapse sequence of the circumgalactic medium driven through feedback \citep{2021MNRAS.501.5176K}. Third and finally, the bimodality is generated through a merger process, either by enhancing the star formation rate (SFR) of the Galaxy \citep{2004ApJ...612..894B,2005ApJ...630..298B,2007ApJ...658...60B,2010MNRAS.402.1489R} or by supplying a relatively pristine gas supply that resets the metallicity of the Galaxy \citep{2020MNRAS.491.5435B,2024MNRAS.528L.122C}. Strong evidence that the Milky Way did undergo a merger with the Gaia-Sausage-Enceladus (GSE) satellite supports the merger-related scenarios \citep{2018MNRAS.478..611B,2018Natur.563...85H,2020ApJ...901...48N,2024ApJ...972..112C}.

@@

In Paper~I, we used idealized simulations of a galaxy merger that triggered the quiescent period. However, in that work we argued the merger was unnecessary. In this work, we study a subhalo, or galaxy, from the Illustris TNG50 simulation demonstrating this. This galaxy exhibits the sequence of events presented in Paper~I after a simple post-processing step which increases the \alphaFe{} of old star particles. This post-processing is motivated by recent work demonstrating that the star formation efficiency (SFE) of the dense gas present at high-$z$ is too low in the TNG model \citep{2024arXiv240909121H}. The galaxy undergoes a brief quiescent period which neatly separates a high- and low-$\alpha$ sequence. However, instead of being preceded by a merger, the quenching is preceded by apparent bar-induced AGN activity. Therefore, this work serves as a verification that the scenario in Paper~I is possible in cosmological simulations and does not require a merger.

In Section~\ref{sec:methods}, we discuss our selection technique which led to discovering the galaxy, the observations we use for comparison, as well as a simple one zone chemical evolution model we use to justify our post-processing step. In Section~\ref{sec:results}, we present the main results which we discuss and interpret in Section~\ref{sec:disc}. We conclude in Section~\ref{sec:conc}.

@@

We select subhalo 172175 (the ID at snapshot 40) for its resemblance to the Milky Way. We then studied the main descendant of this subhalo at $z=0$ (subhalo 392276 at snapshot 99). A summary of its key properties is given in Table~\ref{tab:summ}.

@@

Our galaxy, on the other hand, does not show a clearly bimodal structure in the fiducial simulation (middle column). There is some structure in the $\FeH=-0.75$ bin. The right panel of Figure~\ref{fig:fig1} shows the same distribution as in the middle panel, but with a modification to increase \MgFe{} values of older stars formed before $z=1.5$ (see Sections~\ref{ssec:tng} and \ref{ssec:onezone}). A multimodal structure emerges with three clear modes at $\MgFe\sim0.8$, $0.5$, and $0.2\,\dex$. The 1D histograms show that the modes are well-separated, and that the troughs between the modes nearly vanish.

@@

\begin{figure*}
  \centering
  \includegraphics[width=\textwidth]{fig/392276_alpha.pdf}
  \caption{\textbf{Bimodality in the abundance plane is linked to distinct epochs in simulation.} The upper panels show \MgFe{} as a function of age for our subhalo in TNG. The colors indicate stellar populations at fixed values of \FeH{}, which are the same as in Figure~\ref{fig:fig1}. A gap in the relation occurs at an age of approximately $10.6\,\Gyr$, which we indicate with a vertical dashed line. The effect of the $\alpha$-enhancement is clear, as it separates the stars that form before and after this gap in ages (star particles which formed before $z=1.5$ are $\alpha$-enhanced, which occurs at an age of $\sim9.5\,\Gyr$). The lower panels show on the left the Milky Way and on the center and right the data from TNG but with $12.5\%$ age errors and $0.01\,\dex$ errors in \MgFe{}. When the simulations are given these errors, we see that the before and after star particles smear such that the two populations significantly overlap in ages. In the $\alpha$-enhanced galaxy, two populations emerge in each bin which overlapped in the fiducial distribution. This feature more closely resembles the Milky Way, which displays such populations where the bimodality is strongest -- $\FeH=-0.5$ (blue) and $-0.25$ (orange).}
  \label{fig:alpha}
\end{figure*}

The abundance distributions shown in Figure~\ref{fig:fig1} can be better understood by examining the evolution of \MgFe{} with time of the individual stars/star particles. In the upper panels of Figure~\ref{fig:alpha} we show the true distribution of \MgFe{} as a function of time for the fiducial galaxy in the middle and for the post-processed, $\alpha$-enhanced subhalo to the right. We use age instead of formation time in order to better facilitate comparisons to observations. These panels show 2D histograms, with a logarithmic colormap normalized to the maximum of the plot. To prevent overlap, the values of \MgFe{} are given offsets of $0$, $-0.5$, $-1$, and $-1.5$, in order of increasing \FeH{}.

In our galaxy, there is an age gap at $\sim10.6\,\Gyr$, which we mark with a vertical dashed line in the upper row. Star particles older than this line have a much clearer gradient in time with \MgFe{} than stars that form after. In the $\FeH=-0.25$ bin, star particles which form directly after this line have a slightly reduced \MgFe{} than stars which form a short time later.

In the lower row, the left panel shows data from the Milky Way, while the middle and right panels correspond to the fiducial and $\alpha$-enhanced galaxy, respectively. For the simulations, we introduce Gaussian errors of $12.5\%$ in age and $0.01,\dex$ in \MgFe{}. These error values align with the observational uncertainties from the APOKASC-3 and APOGEE datasets, which the Milky Way panel (lower left) is based on (see Appendix~\ref{app:obs_err} for detailed plots of observational errors). While the error in \MgFe{} is insignificant, the age error ($1.5\,\Gyr$ at $10\,\Gyr$) blurs the distribution, particularly around the dashed line that marks the transition between sequences. Despite this, the $\alpha$-enhanced galaxy still shows two distinct populations, although their ages now overlap more significantly.

Interestingly, the Milky Way distribution (lower left panel) exhibits similarities to the $\alpha$-enhanced galaxy. Specifically, in the $\FeH=-0.5$ and $-0.25$ bins (blue and orange, respectively), there appear to be two populations that overlap in age but remain distinct in \MgFe{}. These are the bins where the bimodality is most pronounced (as seen in the upper left panel of Figure~\ref{fig:fig1}).

@@

To understand the key events driving the behavior around the dashed line in Figure~\ref{fig:alpha}, we examine the evolution of several properties of our galaxy in Figure~\ref{fig:history}. The vertical dashed line in each panel marks the transition at $t\sim3.2\,\Gyr$ from the high- to low-$\alpha$ sequences, as in Figure~\ref{fig:alpha}. The upper panel in the left column shows the the SFR (computed for all gas cells in the galaxy). There are two peaks at $t\sim2.5\,\Gyr$ and $t\sim4.5\,\Gyr$ with maximum values of $50\,\Msun/\textrm{yr}$ and $30\,\Msun/\textrm{yr}$, respectively. Around the high- to low-$\alpha$ transition, there is a dip in the SFR, which drops by an order of magnitude to about $3\,\Msun/\textrm{yr}$. The right panel zooms in on the period between $t=2\,\Gyr$ and $5\,\Gyr$, where we observe a sharp recovery in the SFR following the quiescent phase. In the span of a single snapshot (roughly $150\,\Myr$), the SFR increases from about $3\,\Msun/\textrm{yr}$ to $10\,\Msun/\textrm{yr}$.

The middle panels track the accretion rate of the central BH as a fraction of the Eddington rate. Early in our galaxy's history ($t<2\,\Gyr$), the BH experiences high accretion, which steadily declines until $t\sim5\,\Gyr$. Around $t\sim3.2\,\Gyr$, the BH accretion rate peaks again, reaching approximately $30\%$ of the Eddington limit, placing the BH in quasar mode and injecting significant thermal energy into the galaxy's center. The middle right panel shows the period between $t=2$ and $5\,\Gyr$.

The lower panel illustrates the growth of the galaxy's virial mass ($M_{200}$). Early on ($t<4\,\Gyr$), $M_{200}$ increases roughly linearly, reaching about $2 \times 10^{12}\,\Msun$. After this, the mass remains relatively stable until jumps occur around $t\sim10$ and $\sim12\,\Gyr$, indicative of mergers. There are no mergers related to the earlier quiescent period around $t\sim3.2\,\Gyr$, as no major mass jumps are observed during this time. The lower right panel shows the period between $t=2$ and $5\,\Gyr$.

\subsection{Sequence of Events}\label{ssec:sequence_of_events}
\begin{figure*}
  \centering
  \includegraphics[width=\textwidth]{fig/alyssa.pdf}
  \caption{Surface density projections of gas (top row) and star particles (middle row) in our galaxy across snapshots at different times during the high- to low-$\alpha$ transition. Below the projections is a plot showing the SFR, BH accretion rate (in units of $\dot{M}_{\textrm{BH}}/\dot{M}_{\textrm{edd}}$), and bar strength ($A_2/A_0$ for $R<2$ kpc). Time ranges from $\sim2.4\,\textrm{Gyr}$ to $\sim3.6\,\textrm{Gyr}$, corresponding to redshifts from $z\sim2.7$ to $z\sim1.8$. A sequence of events in which the bar strengthens, BH accretion increases, and SFR declines is seen, and is described more fully in the text.}
  \label{fig:seq}
\end{figure*}

@@

\section{Discussion}\label{sec:disc}
In Figure~\ref{fig:fig1}, we compared the abundance plane between the Milky Way and our galaxy before and after our $\alpha$-enhancement post-processing procedure. It is clear that the TNG galaxy is unimodal before the $\alpha$-enhancement and multi-modal afterwards. Here, we briefly discuss two main points: (1) assuming the $\alpha$-enhancement is justified, what leads to the bimodality in the TNG galaxy?, and (2) what justifies the $\alpha$-enhancement? We then extend our comparison to data and argue the quiescent period is driven by a bar-induced AGN episode.

@@

Once the $\alpha$-enhancement post-processing has been done, the galaxy that we have studied in this work is consistent with this scenario. The vertical dashed line in Figures~\ref{fig:alpha} and \ref{fig:history} denotes the transition between the high- and low-$\alpha$ sequences. We place it at a lookback time of $10.6\,\Gyr$ or a cosmic time of $\sim3.2\,\Gyr$ or $z\sim2.05$.\footnote{The transition between the high- and low-$\alpha$ sequences occurs approximately $1\,\Gyr$ before $z=1.5$, where the $\alpha$-enhancement begins. This, combined with the fact that not all of the subhalos in our sample display bimodalities (Appendix~\ref{app:rand_fig1}), shows that the $\alpha$-enhancement is not the direct cause of the bimodality.} The upper right panel of Figure~\ref{fig:alpha} shows 2D histograms of \MgFe{} vs. star particle age in fixed bins of \FeH{} (each color is a different \FeH{} bin, and an offset is given to \MgFe{} to prevent overlap). Stars formed before the dashed line had a steep decline of \MgFe{} with time, while the relation in star particles formed after the dashed line is much more flat. At the $\alpha$-transition, we see a gap in the $\FeH=-0.5$ and $-0.25$ bins.

@@

It is also worth noting that in both the fiducial and $\alpha$-enhanced galaxy show a slight rebound effect in \MgFe{}. The star particles which form directly after the dashed line have a slightly lower \MgFe{} than stars which form after. This was also predicted in Figure~9 of Paper~I, where it was argued that during the period of suppressed star formation, the \alphaFe{} ratio of the star-forming gas drops sharply due to the reduced contribution of Type~II relative to Type~Ia SNe. Later, the \alphaFe{} of the gas will recover when the SFR also recovers, but there is a brief window when old, low-$\alpha$ stars can form. A similar behavior was seen in the one-zone models with bursty SFHs in \citet{2020MNRAS.498.1364J}.

@@

\subsection{Direct Comparison to Observations}\label{ssec:compare_obs}
The lower left panel of Figure~\ref{fig:alpha} shows the \MgFe{} vs. age of stars in bins of \FeH{} which pass our solar neighborhood selection and are present in the APOKASC-3 catalog. In the bins where the bimodality is strongest (blue and orange, $\FeH=-0.5$ and $-0.25$, respectively), we see that there is a sort of two tiered distribution with significant overlap in age. In the galaxy with age and abundance errors shown in the lower right panel, a similar distribution can be seen in all \FeH{} bins. An examination of the true distribution in the galaxy (upper right panel), we see that the two distributions are in fact very cleanly separated in age.

@@

\subsection{Cause of Quiescence}\label{ssec:cause_qui}
In Paper~I, AGN activity from a merger was the suspected cause for the quiescent period. In our galaxy here, there is indeed a brief burst in AGN accretion at the time of the merger (middle panel of Figure~\ref{fig:history}). Based on this burst, it is also reasonable to suspect that AGN activity is also responsible for the quiescent period in our galaxy. However, we argue that the cause of the localized spike in the BH accretion rate is not due to a merger but instead due to the formation/strengthening of a bar.

@@

A complicating factor for this picture comes from the high SFR associated with the galaxy. The depletion time ($M_{\textrm{gas}}/\textrm{SFR}$) at the $t=2.84\,\Gyr$ snapshot is $204\,\Myr$, shorter than the time it takes for the SFR to reach its minimum. This implies the possibility of starvation as a quenching mechanism. However, the average depletion time in the preceding 5 snapshots (which are each $\sim150\,\Myr$ apart) is $220\,\Myr$, so clearly the galaxy is accreting high amounts of gas from its environment. A definitive account, difficult with the current simulation outputs because of its sparse snapshot spacing, requires further work.

\section{Conclusions}\label{sec:conc}
In this work, we examined a subhalo of interest in Illustris TNG50. This galaxy is at a Milky Way-progenitor mass at $z=1.5$. After applying a post-processing step that increased the \MgFe{} of star particles formed before $z=1.5$, this subhalo hosts a strong bimodality in the plane of \MgFe{} and \FeH{}, shown in Figure~\ref{fig:fig1}. This post-processing is justified by arguing that the SFE of dense gas is too low in TNG \citep[][see discussion in our Section~\ref{ssec:sfe}]{2024arXiv240909121H}.

@@

\section{Random Selection of Subhalos}\label{app:rand_fig1}
In Figures~\ref{fig:app0} to \ref{fig:app16}, we show a version of Figure~\ref{fig:fig1} but for a random selection of subhalos from our broader sample of 168 Milky Way-progenitors at $z=1.5$ in TNG50. It is from this sample that we selected our galaxy of interest. We first show this galaxy at $z=1.5$ in Figure~\ref{fig:app0}, and then show our random sample of 16.

\begin{figure*}
  \centering
  \includegraphics[width=\textwidth]{fig/app_172175.pdf}
  \caption{The same as Figure~\ref{fig:fig1}, but for our galaxy of interest at $z=1.5$.}
  \label{fig:app0}
\end{figure*}
